---
title: "p8105_hw3_yy3307"
output: github_document
date: "2022-10-07"
author: "Yang Yi"
UNI: "yy3307"
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1
The first step we need to do is importing the `instacart` data from `p8105.datasets`.
```{r import instacart}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

_The `Instacart` dataset contains data of online grovery services which allow peopleto shop online from local stores. The dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns, which includes variables of `order_id, product_id, add_to_cart_order, reordered, user_id, eval_set, order_number, order_dow, order_hour_of_day, days_since_prior_order, product_name, aisle_id, department_id, aisle, department`. We can make some data analysis with this dataset, for instance, we can indicate which type of grocery is most popular by counting frequencies of `aisle` variable. Also, we can observe something like the `reorder rate`, the `reorder rate` of this dataset is `r colMeans(instacart[c("reordered")])`, which means that about 59.86% of all customers would like to reorder._

We can use `count` and `distinct` functions to find the total number of aisles. And we also need to return the `name` of aisle which most repeated being ordered, `which.max` can help us find that aisle.

```{r aisles}
num_aisle = count(distinct(instacart, aisle))

aisle_most = names(which.max(table(instacart$aisle)))

except_fresh =
  instacart %>%
  filter(!grepl('fresh vegetables', aisle))
aisle_2most = names(which.max(table(except_fresh$aisle)))
```

_There are **`r num_aisle`** distinct aisles, **`r aisle_most`** and **`r aisle_2most`** are the most items ordered from._

